---
title: "For_publication_intros"
author: "Shadi"
date: "1/23/2022"
output: html_document
---
## This script is a modified version of what we submitted to cogsci. 
## It makes sure that the transition matrices also include the transitions from and to the introduction moments. 
## It also includes a sliding window analysis that replicates the increase in surprisal. 
## This script was used to validate findings from our minimal model of mathematical insights.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(lmerTest)
library(tidygraph)
library(ggraph)
library(gridExtra)
library(igraph)
library(magrittr) 
library(entropy)
library(devtools)
library(ccber) #for probability transition matrix
library(KernSmooth)

```

Here we use the updated list of insights without removing any of the insights that are close to each other
```{r}
insight_data = read_csv(file="insight_math_UPDATED.csv")
```

```{r}

experts <- read.table(file = "mathExperts_allData.txt", 
                      sep='\t', 
                      skip = 1)
colnames(experts) <- c("start","end","duration","ob2","ob1","file")

# xtabs(~(ob1=="") + (ob2==""), experts)
experts <- subset(experts, !(ob1!="" & ob2!=""))
experts <- subset(experts, !(ob1=="" & ob2==""))

experts <- experts %>% 
  separate(file, into = c("subName","problemName","elan"), remove = T) %>% 
 dplyr:: mutate(ob = paste(ob1, ob2, sep=""),
         ob = as.integer(as.character(gsub("Object| ","", ob))),
         subProb = paste(subName, problemName, sep="-")) %>% 
  dplyr::select(subName, problemName, start, end, duration, ob, subProb)

experts <- experts %>% 
  filter(subProb != "Prasit-PowerSet") #only person who did this problem



```

```{r}

#Making new labels for objects (basically trying to reduce the number of individual objects originally determined by Elan program).
newObjectLabels <- read.csv(file = "Experts-RecodingObjectsIntoClusters.csv")

colnames(newObjectLabels) <- c("subName","problemNum","ob","obNew","problemName")

experts.2 <- experts %>% 
  left_join(newObjectLabels, by=c("subName", "problemName","ob")) %>% 
  mutate(obNew = ifelse(is.na(obNew), ob, obNew)) %>% 
  dplyr::group_by(subName, problemName) %>% 
  mutate(obNew = as.integer(as.factor(obNew))) %>% 
  dplyr::group_by(subProb) %>% 
  mutate(totalObjects = max(obNew),
         startn = start/max(end),
         durationn = duration/max(end),
         endn = end/max(end)) %>%
  ungroup() 



# Selecting the relevant columns from the original insight data frame.
adapted_insight<- insight_data %>% 
  filter(keep == T) %>% 
  dplyr::select(subName,problemName,start, Event, Insight_Number,insight_intensity )


# Adding a new column to specify the type of event (i.e., attention or insight).
experts.3 <- experts.2 %>%
  mutate(Event= "attention")

# Adding rows of the insight data frame to the original data frame including information about participants' interactions with objects (i.e., mathematical notations).
experts_and_insights <- bind_rows(experts.3,adapted_insight )

experts_and_insights_clean<-experts_and_insights %>% 
  filter(subName!= "Neil", subName!= "Prasit")%>% # Neil only did one of the problems, and no information for Prasit's insights is available. So they are excluded.
  dplyr::group_by(subName, problemName)%>%
  arrange(subName, problemName, start) %>%
  mutate(iei = c(NA, diff(start)))

```

```{r}
index <- which(experts_and_insights_clean==1726070, arr.ind=TRUE)
index # row 2831 and row 2832 have the same start time which causes problems in later analyses. So I will add jitter to one of those lines.
#That is I add 20ms to the start value on row 2832

library(magrittr)
experts_and_insights_clean<-experts_and_insights_clean %>% { .[2832,3] <- 1726090;. }
unique(experts_and_insights_clean$totalObjects)
```

#make a for loop to count the number of rows before the first insight;

```{r}

count_row_prior_1st_insight <- tibble()
for (subject in unique(experts_and_insights_clean$subName)){ # Subsetting by participants 
  subset_subj <- experts_and_insights_clean %>%
    filter(subName == subject) 
  
  for (problem in unique(subset_subj$problemName)){ # Subsetting by problems
    subset_prob <- subset_subj %>%
      filter(problemName == problem) %>%
      mutate(num_prior_event = NULL) 
    
    for (i in length(subset_prob$Event)){
      if (i == "insight"){num_prior_event = length(1:i)
      } else {
        next
      }
      
      add<- tibble (start = subset_prob$start, subject = subject, problem = problem,num_prior_event = num_prior_event )
      
      count_row_prior_1st_insight <- bind_rows(count_row_prior_1st_insight, add)
    }
  }
}


```

# Calculating transition matrix both based on all events and prior events
```{r}
# Making a function to calculate entropy from transition matrix
ent_from_trans_matrix <- function(trans_matrix, val_prev, val_next){
  prob = trans_matrix[val_prev, val_next]
  h = -log2(prob)
  return(h)
}

```

```{r}
# This code calculates the transition matrix based on all events in each session, 
# it also calculates the transition matrix based only on prior events.

#pay attention that we filter for attention to get rid of the insight moment since they are npt associated with an inscription. Intro moments are not marked in this dataset (experts_and_insights_clean).

trans_cont_time_ent <- tibble()
window_size <- 60*1000
for (subject in unique(experts_and_insights_clean$subName)){ # Subsetting by participants 
  subset_subj <- experts_and_insights_clean %>%
    filter(subName == subject) %>%
    filter(Event == "attention")
  
  for (problem in unique(subset_subj$problemName)){ # Subsetting by problems
    subset_prob <- subset_subj %>%
      filter(problemName == problem) %>%
      # Making a new column including a lagged version of the objects:
      mutate(obNew_fac = factor(obNew),
             obNew_fac_prev = lag(obNew_fac)) 
    
    subset_prob_with_intros <- experts_and_insights_clean %>%
      filter(subName == subject, problemName == problem) %>%
      filter(Event != "insight") %>%
      mutate(obNew_fac = factor(obNew),
             obNew_fac_prev = lag(obNew_fac)) 
    
    #### Making a matrix of all transition counts:
    trans_count <- CalcTransitionCounts(subset_prob_with_intros$obNew_fac)
    # Calculating the probability of each transition:
    trans_prob <- CalcTransitionMatrix(trans_count) 
    
    for(eventi in 2:nrow(subset_prob)){
      # Subset data so it's only the current event and prior events
      current_event <- subset_prob[eventi,]
      print(paste(current_event$subName, current_event$problemName, eventi))
      event_time <- current_event$start
      
      prior_events <- subset_prob_with_intros %>% 
        filter(start > (event_time - window_size),
               start <= event_time) %>%
        mutate(obNew_fac_prior = factor(obNew),
               obNew_fac_prev_prior = lag(obNew_fac_prior),
               prior_total_obj= n_distinct(obNew)) 
      
      current_event <- subset(prior_events, start == event_time)
      
      # Get transition matrix for those events
      trans_count_prior <- CalcTransitionCounts(prior_events$obNew_fac_prior)
      trans_prob_prior <- CalcTransitionMatrix(trans_count_prior) 
      # Calculate entropy of the current event, relative to prior events
      current_event$current_event_ent_prior <- ent_from_trans_matrix(trans_prob_prior, 
                                                                     current_event$obNew_fac_prev_prior, 
                                                                     current_event$obNew_fac_prior)
      # Calculate entropy of the current event, relative to all events
      current_event$current_event_ent_overall <- ent_from_trans_matrix(trans_prob,current_event$obNew_fac_prev,current_event$obNew_fac)
      
      trans_cont_time_ent <- bind_rows(trans_cont_time_ent,current_event)
    }
  }
}
hist(trans_cont_time_ent$current_event_ent_prior)
```

## Adding introductions to the Event Column:

```{r}
All_measures_and_events <- tibble()

for (subject in unique(trans_cont_time_ent$subName)){ #subset by person
  subset_subj <- trans_cont_time_ent %>% 
    filter(subName == subject, is.na(obNew) == F) #Get rid of the NA rows
  print(subject)
  for (problem in unique(subset_subj$problemName)){ #subset by problem
    print(problem)
    subset_prob <- subset_subj %>% 
      filter(problemName == problem)
    unique_objects <- sort(unique(subset_prob$obNew))
    for (i in 1:nrow(subset_prob)) { # for each row in the problem
      if (subset_prob$obNew[i] %in% unique_objects == TRUE) { # if the object is one found in the unique_objects vector
        print(paste0(i))
        #unique_objects <- unique_objects[-subset_prob$obNew[i]] # remove the objects number from the vector
        unique_objects <- unique_objects[unique_objects != subset_prob$obNew[i] ]
        subset_prob$Event[i] <- "intro" # change the event for that row to "intro"
      } else {
        next
      }
    }
    All_measures_and_events <- bind_rows(All_measures_and_events, subset_prob)
  }
}


```



```{r}

#Marking the relation of shifts of attention to the moment of insight
window_param <- 60 * 1000

All_measures_and_events$isAfterInsight <- F
All_measures_and_events$isBeforeInsight <- F

All_measures_and_events <- All_measures_and_events %>%
  mutate(event10 = (Event == "intro"),
         startMinutes = start / (60*1000),
         isAroundInsight = isAfterInsight | isBeforeInsight)

for(insight_i in 1:nrow(adapted_insight)){
  # print(insight_i)
  #get insight info
  subNamei <- as.character(adapted_insight[insight_i,"subName"])
  problemNamei <- as.character(adapted_insight[insight_i,"problemName"])
  starti <- as.numeric(adapted_insight[insight_i,"start"])
  
  All_measures_and_events[All_measures_and_events$subName == subNamei &
                            All_measures_and_events$problemName == problemNamei & 
                            All_measures_and_events$start >= (starti - window_param) & 
                            (All_measures_and_events$start < starti),]$isBeforeInsight <- T
  
  All_measures_and_events[All_measures_and_events$subName == subNamei &
                            All_measures_and_events$problemName == problemNamei & 
                            (All_measures_and_events$start > starti) & 
                            All_measures_and_events$start <= (starti + window_param),]$isAfterInsight <- T
  
}

xtabs(~isBeforeInsight + isAfterInsight, All_measures_and_events)

```

```{r}

All_measures_and_events.ml1 <- glmer(event10 ~ 1 + isBeforeInsight * isAfterInsight + startMinutes + 
                                       (1|subProb), 
                                     data = All_measures_and_events, 
                                     family = "binomial", 
                                     verbose=T)
summary(All_measures_and_events.ml1)

All_measures_and_events.ml2 <- lmer(current_event_ent_prior ~ 1 + event10 + 
                                      isBeforeInsight * isAfterInsight + 
                                      startMinutes + 
                                      (1|subProb), 
                                    data = All_measures_and_events)
summary(All_measures_and_events.ml2)
```

```{r,fig.width = 8.5, fig.height = 4, warning=FALSE}
#correct plot codes:
All_measures_and_events <- All_measures_and_events %>% 
  mutate(relToInsight = case_when(isBeforeInsight & !isAfterInsight ~ "pre-insight", 
                                  isAfterInsight & !isBeforeInsight ~ "post-insight", 
                                  isAfterInsight & isAfterInsight ~ "both", 
                                  TRUE ~ "non-insight")) 

All_measures_and_events.agg <- All_measures_and_events %>%
  group_by(relToInsight, isBeforeInsight, isAfterInsight, subProb) %>%
  dplyr::summarize(new_objects_M = mean(event10),
                   ent_M = mean(current_event_ent_prior),
                   .groups = "drop") %>%
  group_by(relToInsight, isBeforeInsight, isAfterInsight) %>%
  dplyr::summarize(new_objects_MM = mean(new_objects_M),
                   new_objects_SE = sd(new_objects_M)/sqrt(n()),
                   ent_MM = mean(ent_M),
                   ent_SE = sd(ent_M)/sqrt(n()),
                   .groups = "drop")

All_measures_and_events.agg <- All_measures_and_events.agg %>% 
  mutate(plot_x = case_when(relToInsight == "non-insight" ~ 1, 
                            relToInsight ==  "pre-insight"  ~ 3, 
                            relToInsight == "post-insight" ~ 4), 
         illumination = case_when(relToInsight == "non-insight" ~ F, 
                                  relToInsight ==  "pre-insight"  ~ T, 
                                  relToInsight == "post-insight" ~ T))

measures_dataframe <- data.frame(All_measures_and_events.agg)

measures_dataframe %>% 
  ggplot(aes(x= plot_x, y = ent_MM, color = illumination)) + 
  geom_pointrange(aes(ymin = ent_MM - ent_SE, 
                      ymax = ent_MM + ent_SE), size = 2) + 
  ylab("surprisal") +
  scale_x_continuous( label = c( "non-insight", "pre-insight", "post-insight"), breaks = c(1,3,4), limits = c(0,4.5)) +
  scale_y_continuous (expand = c(0,0.01), limits = c(1, 1.4))+
  theme_classic(base_size = 30)+
  theme(axis.text = element_text(size = 40)) +
  theme(text = element_text(size = 40)) +
  scale_color_discrete(guide=F) +
  theme(axis.title.x = element_blank())+
  geom_vline(xintercept = 2, linetype = 2)

```

```{r,fig.width = 10, fig.height = 6 }

#surprisal relative to the moment of insight
actual_surp <- measures_dataframe %>% 
  # filter(plot_x != "both") %>%
  ggplot(aes(x= plot_x, y = ent_MM, color = illumination)) + 
  geom_pointrange(aes(ymin = ent_MM - ent_SE, 
                      ymax = ent_MM + ent_SE), size = 2) + 
  ylab("surprisal") +
  scale_x_continuous( label = c( "non\ninsight", "pre\ninsight", "post\ninsight"), breaks = c(1,3,4), limits = c(0,4.5)) +
  scale_y_continuous (expand = c(0,0.01), limits = c(1, 1.4))+
  theme_classic(base_size = 30)+
  theme(axis.text = element_text(size = 30)) +
  theme(text = element_text(size = 30)) +
  theme(axis.line = element_line(size = 1))+
  scale_color_discrete(guide=F) +
  theme(axis.title.x = element_blank())+
  geom_vline(xintercept = 2, linetype = 2, size = 1)



```


# Sliding window analysis of surprisal
```{r}

sliding_width <- 1*60*1000
sliding_step <- 1000
sliding_window_df <- tibble()
for (subProbi in unique(All_measures_and_events$subProb)){
  print(subProbi)
  All_measures_and_events.sub <- All_measures_and_events %>%
    filter(subProb == subProbi)
  for(starti in seq(sliding_width, 
                    ceiling(max(All_measures_and_events.sub$start)/sliding_step)*sliding_step, 
                    sliding_step)){
    windowed_i <- All_measures_and_events.sub %>%
      filter(start > (starti - sliding_width), start <= starti) %>%
      group_by(subName, problemName, subProb) %>%
      dplyr::summarize(surprisalM = mean(current_event_ent_prior),
                       start = starti,
                       n = n(),
                       .groups="drop")
    sliding_window_df <- bind_rows(sliding_window_df, windowed_i)
  }
}
```


```{r, fig.width = 10, fig.height = 6}
# Pulling out relevant periods and recentering the time variable so it's centered on the first insight:
adapted_insight <- adapted_insight %>%
  mutate(subProb = paste(subName, problemName, sep="-"))
first_last_insights <- adapted_insight %>%
  dplyr::group_by(subProb) %>%
  dplyr:: summarise(first_insight_start = round(min(start)/sliding_step)*sliding_step,
                    last_insight_start = round(max(start)/sliding_step)*sliding_step,
                    .groups="drop")
sliding_window_df2 <- sliding_window_df %>%
  left_join(first_last_insights)
sliding_window_df2 <- sliding_window_df2 %>%
  mutate(start_first_c = start - first_insight_start,
         start_last_c = start - last_insight_start)
time_before <- 5*60*1000
time_after <- 3*60*1000
sliding_window_df.insightPeriods <- sliding_window_df2 %>%
  filter(start_first_c > -time_before,
         start_first_c < time_after)


sliding_window_df.insightPeriods.agg <- sliding_window_df.insightPeriods %>%
  group_by(start_first_c) %>%
  dplyr::summarize(surprisalM_M = mean(surprisalM),
                   surprisalM_SE = sd(surprisalM)/sqrt(n()),
                   n = n(), .groups = "drop")
start_value <- min(sliding_window_df.insightPeriods.agg$surprisalM_M)
start_time <- min(sliding_window_df.insightPeriods.agg$start_first_c)/(60*1000)

actual_slide <- sliding_window_df.insightPeriods.agg %>%
  mutate(start_first_c_min = start_first_c/(60*1000)) %>%
  filter(n>1) %>%
  ggplot(aes(x=start_first_c_min, y=surprisalM_M)) +
  geom_ribbon(aes(ymin=surprisalM_M - surprisalM_SE, ymax=surprisalM_M + surprisalM_SE), alpha=.5) +
  geom_line(size = 1) +
  geom_vline(xintercept=0, size = 1) +
  geom_segment(aes(x = start_time - 1, y = start_value, xend = start_time, yend = start_value), size = 1, colour = "blue")+
  theme_classic(base_size = 30)+
  theme(axis.text = element_text(size = 25)) +
  theme(text = element_text(size = 30)) +
  theme(axis.line = element_line(size = 1))+
  scale_color_discrete(guide=F) +
  scale_x_continuous(expression(paste("time (", italic("t"), ")"))) +
  scale_y_continuous(expression(paste("surprisal (", italic("h(t)"), ")")))


```


```{r,  fig.width = 22, fig.height = 6, warning=FALSE}
library(cowplot)
plot_grid(actual_surp, actual_slide, ncol = 2, nrow = 1, align= "hv", labels= NULL)
```



