---
title: "mini_model_fixed_gamma_sep_7_24"
author: "Shadi"
date: '2022-11-09'
output: html_document
---
In this script, we run the model using fix values of gamma (the control variable) to test model results without conflating time trend effects.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(lmerTest)
library(tidygraph)
library(ccber) #for probability transition matrix
library(rootSolve)
library(cowplot)

```


```{r}

solve_sde <- function(
    f, beta, x0, gamma0, delta_t = 0.01,
    nr_iter = 1000, sigma = 0.001,
    initial_period = 200
) {
  
  xt <- rep(x0, nr_iter)
  gammat <- rep(gamma0, nr_iter)
  
  for (n in seq(2, nr_iter + 1)) {
    
    if(n < initial_period){ # Initial period when they're stuck in non-insight
      gammat[n] <- gamma0
    } else { # After initial period, gradually changing the control parameter gamma to encourage insight
      gammat[n] <- gammat[n-1] + delta_t * beta
    }
    
    xt[n] <- xt[n-1] + delta_t * f(xt[n-1], gammat[n-1]) + rnorm(1, 0, sigma)
    if(xt[n] <= 0) xt[n] <- 0.01 # keep value above zero
    if(xt[n] >= 1) xt[n] <- 0.99 # keep value above zero
  }
  
  time <- seq(0, nr_iter * delta_t, delta_t)
  res <- cbind(time, xt, gammat)
  res
}



insight_ds <- function(x, gamma) gamma * (1 - x) ^2 / (0.10^2 + (1 - x)^2) - x * (1 - x)

x <- seq(0, 1, 0.01)
gammas <- seq(0.18, 0.26, 0.01)

```

# For one value of gamma
```{r}
delta_t <- 0.01
nr_iter <- 100000
#set.seed(1)

# set x0 to the attractor in the relevant basin
#wrap this code in a for loop that goes over values of gamma, do one run for each and look at sd and ac / data frame includes gamma, t, (15000 rows for each gamma), for each gamma create a transition matrix and surprial and the mean surprisal for each value of gamma.
df <- function(x, gamma=.19) (gamma) * ((1-x)^2 / ((1-x)^2+(.1)^2) ) - x * (1-x)

X0_attractor1 <- uniroot.all(f = df, interval=c(0.001,0.999))[1]
X0_attractor2 <- uniroot.all(f = df, interval=c(0.001,0.999))[3]
X0_unstable <- uniroot.all(f = df, interval=c(0.001,0.999))[2]


res1 <- solve_sde(
  f = insight_ds, beta = 0, x0 = X0_attractor1, gamma0 = .19,
  nr_iter = nr_iter, delta_t = delta_t, sigma = 0.003, initial_period = 0
)

res2 <- solve_sde(
  f = insight_ds, beta = 0, x0 = X0_attractor2, gamma0 = .19,
  nr_iter = nr_iter, delta_t = delta_t, sigma = 0.003, initial_period = 0
)

res3 <- solve_sde(
  f = insight_ds, beta = 0, x0 = X0_unstable, gamma0 = .19,
  nr_iter = nr_iter, delta_t = delta_t, sigma = 0.003, initial_period = 0
)
```


# for loop attempt
```{r}
delta_t <- 0.01
nr_iter <- 100000
set.seed(1)

# set x0 to the attractor in the relevant basin
#wrap this code in a for loop that goes over values of gamma, do one run for each and look at sd and ac / data frame includes gamma, t, (15000 rows for each gamma), for each gamma create a transition matrix and surprisal and the mean surprisal for each value of gamma.

#df <- function(x, gamma=.19) (gamma) * ((1-x)^2 / ((1-x)^2+(.1)^2) ) - x * (1-x)

all_gamma <- tibble()
for (i in gammas){
  df <- function(x, gamma=i) (gamma) * ((1-x)^2 / ((1-x)^2+(.1)^2) ) - x * (1-x)
  
  X0_attractor1 <- uniroot.all(f = df, interval=c(0.001,0.999))[1]
  X0_attractor2 <- uniroot.all(f = df, interval=c(0.001,0.999))[3]
  X0_unstable <- uniroot.all(f = df, interval=c(0.001,0.999))[2]
  
  res1 <- solve_sde(
    f = insight_ds, beta = 0, x0 = X0_attractor1, gamma0 = i,
    nr_iter = nr_iter, delta_t = delta_t, sigma = 0.003, initial_period = 0
  )
  
  res2 <- solve_sde(
    f = insight_ds, beta = 0, x0 = X0_attractor2, gamma0 = i,
    nr_iter = nr_iter, delta_t = delta_t, sigma = 0.003, initial_period = 0
  )
  
  res3 <- solve_sde(
    f = insight_ds, beta = 0, x0 = X0_unstable, gamma0 = i,
    nr_iter = nr_iter, delta_t = delta_t, sigma = 0.003, initial_period = 0
  )
  
  SD_att_1 <- sd(res1)
  SD_att_2 <- sd(res2)
  SD_att_unst <- sd(res3)
  
  ac_att_1 <- acf(res1, lag.max=10, plot = F)$acf[10]
  ac_att_2 <- acf(res2, lag.max=10, plot = F)$acf[10]
  acc_att_unst <-acf(res3, lag.max=10, plot = F)$acf[10]
  
  add <- tibble(SD_att_1, SD_att_2, SD_att_unst,ac_att_1,ac_att_2, acc_att_unst, gamma = i )
  all_gamma <- rbind(add, all_gamma)
  
}

```
## Just one starting point


```{r}
gammas <- seq(.1, .3, .01)
nr_iter <- 15000
run_num <- 10

all_gamma_sameX0 <- tibble()
for(run_i in 1:run_num){
  print(run_i)
  for (i in gammas){
    df <- function(x, gamma=i) (gamma) * ((1-x)^2 / ((1-x)^2+(.1)^2) ) - x * (1-x)
    
    X0 <- uniroot.all(f = df, interval=c(0.001,0.999))[1]
    
    res1 <- solve_sde(
      f = insight_ds, beta = 0, x0 = X0, gamma0 = i,
      nr_iter = nr_iter, delta_t = delta_t, sigma = 0.003, initial_period = 0
    )
    
    add <- tibble(time = res1[,"time"], U = res1[,"xt"],  gammat = res1[,"gammat"], X0 = X0, run = run_i) %>%
      mutate(gamma_run = paste(gammat, run, sep = "-"))
    all_gamma_sameX0 <- rbind(all_gamma_sameX0, add)
    
  }
}

```

```{r}
# Making a random matrix

make_random_prob_matrix <- function(n) {
  
  all_probabilities <- matrix(nrow = n, ncol = n)
  
  for (i in 1:n) {
    probability_vector <- runif(1, 0, 1)
    for (j in 2:n-1) {
      probability_vector[j] <- runif(1, 0 , (1 - (sum(probability_vector[1:j-1], na.rm = TRUE))))
    }
    probability_vector[n] <- 1 - sum(probability_vector)
    
    # message(sum(probability_vector))
    probability_vector_shuffle <- sample(probability_vector, replace = F)
    
    all_probabilities[i, ] <- probability_vector_shuffle
  }
  return(all_probabilities)
}

``` 

```{r}

U_multiple_fixed_gam <- all_gamma_sameX0

n_ob <- 25
object_list <- c(1:n_ob) # a list of objects

created_objects <- vector() # an empty vector for objects that will be introduced
all_objects_fixed_gam <- tibble()
current_object <- 1  # introducing the very first object
t_since_new_intro <- 0 #time since a new object was introduced


for (gamma_run_i in unique(U_multiple_fixed_gam$gamma_run)){ #subset by person
  print(gamma_run_i)
  p_matrix_noninsight <- make_random_prob_matrix(n_ob)  # random probability matrix
  p_matrix_insight <- make_random_prob_matrix(n_ob)  # random probability matrix
  
  subset_subj <- U_multiple_fixed_gam %>% 
    filter(gamma_run == gamma_run_i) %>% #Get rid of the NA rows
    ungroup() %>% 
    mutate(rowNum = 1:n()) %>%
    filter(rowNum %% 30 == 0) # Thin out the dataset, so an "object interaction" event only happens every couple timesteps.
  
  created_objects <- vector()  #reset values for each subject
  t_since_new_intro <- 0  #reset values for each subject
  
  
  for(i in 1:nrow(subset_subj)){ 
    U_current <- subset_subj[[i, "U"]] 
    U_current <- case_when(U_current <= 1 & U_current >= 0 ~ U_current, #making sure that the U values fall between zero and one
                           U_current > 1 ~ 1, 
                           TRUE ~ 0)
    p_matrix_current <- (1-U_current) * p_matrix_noninsight + (U_current) * p_matrix_insight #using U values to find the network of concepts at each time (all the way from the non-insight to insight)
    
    trans_probs_justCurrentObject <- p_matrix_current[current_object, ] #getting transition probability
    
    
    next_object <- sample(object_list, 1, prob = trans_probs_justCurrentObject) # update the new object by randomly choosing an object from the object list based on the probability matrix.
    
    
    if (next_object %in% created_objects) {  # mark when the object are introduced ("intro") and when they are interacted with ("attention")
      event <- "attention"
      t_since_new_intro <- t_since_new_intro + 1 # determine how many time steps passed from the last intro event.
    } else {
      created_objects <- c(created_objects, next_object)
      event <- "intro"
      t_since_new_intro <- 0 # obviously when an event is intro, there is no time steps between the event and itself, therefore 0.
    }
    add <- tibble(gamma_run = gamma_run_i,
                  i = i, 
                  time = subset_subj[[i, "time"]], U = U_current, 
                  gamma = subset_subj[[i, "gammat"]],
                  run = subset_subj[[i, "run"]],
                  ob = current_object, 
                  ob_next = next_object, event = event, t_since_new_intro)
    all_objects_fixed_gam <- bind_rows(all_objects_fixed_gam, add)
    current_object <- next_object
  }
}


```




# add surprisal for each gamma set

```{r}
# Making a function to calculate entropy from transition matrix
ent_from_trans_matrix <- function(trans_matrix, val_prev, val_next){
  prob = trans_matrix[val_prev, val_next]
  h = -log2(prob)
  return(h)
}

```

```{r}

# Time to calculate surprisal.

add_surprisal_fixed_gam <- tibble()

window_size <- 12 

for(gamma_run_i in unique(all_objects_fixed_gam$gamma_run)){
  print(gamma_run_i)
  session_filter <- all_objects_fixed_gam %>%
    filter(gamma_run == gamma_run_i) %>%
    filter (event == "attention") %>%
    mutate(obNew_fac = factor(ob), # make "ob" column into factor.
           obNew_fac_prev = lag(ob)) # add a lagged version of "ob" column.
  
  subset_prob_with_intros <- all_objects_fixed_gam %>%
    filter(gamma_run == gamma_run_i) %>%
    mutate(obNew_fac = factor(ob),
           obNew_fac_prev = lag(obNew_fac)) 
  
  #### Making a matrix of all transition counts:
  trans_count <- CalcTransitionCounts(subset_prob_with_intros$obNew_fac)
  # Calculating the probability of each transition:
  trans_prob <- CalcTransitionMatrix(trans_count) 
  
  # Find out which row to start the for loop at — we need enough data to fill the entire sliding window
  start_row <- session_filter %>% 
    mutate(row = 1:n()) %>%
    filter(time < window_size) 
  start_row <- start_row %>% filter(time == max(start_row$time))
  start_row <- start_row$row + 1
  
  for(eventi in start_row:nrow(session_filter)){
    # Subset data so it's only the current event and prior events
    current_event <- session_filter [eventi,]
    event_time <- current_event$time
    
    prior_events <- subset_prob_with_intros %>% 
      filter(time > (event_time - window_size), 
             time <= event_time) %>%
      mutate(obNew_fac_prior = factor(ob),
             obNew_fac_prev_prior = lag(obNew_fac_prior),
             prior_total_obj= n_distinct(ob)) 
    
    current_event <- subset(prior_events, time == event_time)
    
    # Get transition matrix for those events
    trans_count_prior <- CalcTransitionCounts(prior_events$obNew_fac_prior)
    trans_prob_prior <- CalcTransitionMatrix(trans_count_prior) 
    # Calculate surprisal of the current event, relative to prior events
    current_event$current_event_ent_prior <- ent_from_trans_matrix(trans_prob_prior,current_event$obNew_fac_prev_prior,current_event$obNew_fac_prior)
    # Calculate entropy of the current event, relative to all events
    current_event$current_event_ent_overall <- ent_from_trans_matrix(trans_prob,current_event$obNew_fac_prev,current_event$obNew_fac)
    
    add_surprisal_fixed_gam <- bind_rows(add_surprisal_fixed_gam,current_event)
  }
}

#write_csv(add_surprisal_fixed_gam, file = "fixed_gamma_final_plots.csv" )
#add_surprisal <- add_surprisal %>%
# group_by(run, total_obj) %>%
#arrange(by_group = T)

#add_surprisal
```
```{r}
str(add_surprisal_fixed_gam)
str(add_surprisal_fixed_gam$current_event_ent_prior)
```


##point range version of the plots

```{r, fig.width = 8, fig.height = 6, warning=FALSE,}

add_surprisal_fixed_gam<- read.csv(file = "fixed_gamma_final_plots.csv")

add_surprisal_fixed_gam_agg_1 <- add_surprisal_fixed_gam %>%
  group_by(gamma, run) %>%
  summarize(CF = sd(U),
            mean_h = mean(current_event_ent_prior))

add_surprisal_fixed_gam_agg_2 <- add_surprisal_fixed_gam_agg_1 %>% 
  group_by(gamma) %>% 
  summarize(M_CF =mean(CF),
            se_CF = sd(CF)/sqrt(n()),
            MM_h = mean(mean_h),
            se_mean_h = sd(mean_h)/sqrt(n()))




CF <- add_surprisal_fixed_gam_agg_2 %>%
  group_by(gamma) %>%
  filter(gamma < .25) %>%
  ggplot(aes(x= gamma, y = M_CF)) +
  #geom_point(aes(y=), color= "#009E73", size = 1) +
  geom_pointrange(aes(ymin = M_CF - se_CF, 
                      ymax = M_CF + se_CF), color= "#009E73", size = .70) +
  geom_smooth(method="lm", color = "#999999") + 
  ylab("critical fluctuations, \nSD(U)") +
  xlab (expression ("\ndrive to explore" ~ (gamma))) +
  scale_y_continuous(breaks = c(.020, .030, .040))+
  theme_classic(base_size = 30)+
  theme(axis.text = element_text(size = 30)) +
  theme(text = element_text(size = 30)) +
  theme(axis.line = element_line(size = 1))+
  scale_color_discrete(guide=F) 



SURP <- add_surprisal_fixed_gam_agg_2 %>%
  group_by(gamma) %>%
  filter(gamma < .25) %>%
  ggplot(aes(x= gamma, y = MM_h)) +
  #geom_point(aes(y=), color= "#009E73", size = 1) +
  geom_pointrange(aes(ymin = MM_h - se_mean_h, 
                      ymax = MM_h + se_mean_h), color= "#009E73", size = .70) +
  geom_smooth(method="lm", color = "#999999") + 
  ylab("surprisal") +
  xlab (expression ("\ndrive to explore" ~ (gamma))) +
  scale_y_continuous(breaks = c(.75, .85 , .95), limits = c(.74, 1))+
  theme_classic(base_size = 30)+
  theme(axis.text = element_text(size = 30)) +
  theme(text = element_text(size = 30)) +
  theme(axis.line = element_line(size = 1))+
  scale_color_discrete(guide=F) 



# for surprisal on sd

CF_SURP <- add_surprisal_fixed_gam_agg_2 %>%
  #group_by(gamma) %>%
  filter(gamma < .25) %>%
  ggplot(aes(x= M_CF, y = MM_h)) +
  geom_pointrange(aes(ymin = MM_h - se_mean_h, 
                      ymax = MM_h + se_mean_h), color= "#009E73", size = .70) +
  geom_smooth(method="lm", color = "#999999") + 
  ylab("surprisal") +
  xlab ("\ncritical fluctuations, \nSD(U)") +
  scale_y_continuous(breaks = c(.75, .85 , .95),  limits = c(.74, 1))+
  scale_x_continuous(breaks = c(.025, .035, .045))+
  theme_classic(base_size = 30)+
  theme(axis.text = element_text(size = 30)) +
  theme(text = element_text(size = 30)) +
  theme(axis.line = element_line(size = 1))+
  scale_color_discrete(guide=F) 

```

##Pearson correlation results:
#gam_less_25 <- add_surprisal_fixed_gam_agg_2%>% 
  #filter(gamma <.25)
# cor.test(gam_less_25$M_CF, gam_less_25$MM_h)
# 
# Pearson's product-moment correlation
# 
# data:  gam_less_25$M_CF and gam_less_25$MM_h
# t = 5.1545, df = 13, p-value = 0.0001852
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.5293699 0.9379659
# sample estimates:
#       cor 
# 0.8194241 

# gam_less_25 <- add_surprisal_fixed_gam_agg_2%>% 
#   filter(gamma <.25)
# cor.test(gam_less_25$gamma, gam_less_25$M_CF)
# 
# Pearson's product-moment correlation
# 
# data:  gam_less_25$gamma and gam_less_25$M_CF
# t = 5.7634, df = 13, p-value = 6.562e-05
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.5930691 0.9482340
# sample estimates:
#       cor 
# 0.8477725 

# gam_less_25 <- add_surprisal_fixed_gam_agg_2%>% 
#   filter(gamma <.25)
# cor.test(gam_less_25$gamma, gam_less_25$MM_h)
# 
# Pearson's product-moment correlation
# 
# data:  gam_less_25$gamma and gam_less_25$MM_h
# t = 7.4178, df = 13, p-value = 5.063e-06
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.7178486 0.9664040
# sample estimates:
#       cor 
# 0.8993837 



```{r,  fig.width = 20, fig.height = 6, warning=FALSE}
library(cowplot)
plot_grid(CF, SURP, CF_SURP, ncol = 3, nrow = 1, align= "hv", labels= "AUTO", label_x = .1, label_size = 30)
```


```{r}
grid.arrange(CF, SURP, CF_SURP, ncol=3, nrow = 1, newpage = T )
```

